### 2025-09-08 project-plan.Rmd

---
title: "Manuscript title here"
subtitle: "Add a Subtitle if Needed"
author: "Your name here"
bibliography: bibliography.bib
csl: vancouver.csl
knit: (function(inputFile, encoding) {
  rmarkdown::render(inputFile, encoding = encoding, output_dir = "out") })
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<!-- The project plan should be 3-4 pages long and written in
Rmarkdown (like this document). Remove all comments in italic when you 
use this document as a template for your own project plan. -->

<!-- This is an R Markdown document. Markdown is a simple formatting
syntax for authoring HTML, PDF, and MS Word documents. For more
details on using R Markdown see http://rmarkdown.rstudio.com -->

<!-- This is a comment and will not be present in the compiled
document. Try "knitting" this document by clicking the knit button up
to the left, or by running `rmarkdown::render("project-plan.Rmd")` in
the console and you should only see headings -->

Introduction
============

*This is part A) in the project plan [sv: projektredogörelse]. *

*The background/introduction should summarise the scope of the problem,
what is known about the problem, what is not known, what specific
knowledge gap the study is supposed to fill and why this is
important. It should end be stating the aim of the study. The 
background should be 4-5 paragraphs long and each paragraph should be
between 3-5 sentences. Strive to make the paragraphs about the same
length.*

*bibliography.bib includes an example reference. Add additional
references in that file as bibtex entries and then cite in text using:
"[@exampleKey9999]". One can convert DOIs' to bibtex entries using 
https://www.doi2bib.org/. If you use Zotero you can simply export 
entries to bibtex `Ctrl-Shift-C`, however you need to first setup 
Quick Copy to export to bibtex (https://www.zotero.org/support/creating_bibliographies#quick_copy)*

Methods
=======

*Refer to the appropriate reporting guideline for details. If you are
developing, updating or validating a clinical prediction model then
use
[TRIPOD](https://www.equator-network.org/reporting-guidelines/tripod-statement/). If
you are conducting an observational study, for example a cohort or
case control study in which you assess associations between some
exposure and an outcome then use
[STROBE](https://www.equator-network.org/reporting-guidelines/strobe/).*

Project Update and Timeline
===========================

*This is part B) in the project plan [sv: projektredogörelse]. *

*Save the image generated from the Gantt chard spreadsheet into the assets folder, rename it to project-gantt.png*

![**The Gantt chart will be displayed here when you place it in the assets folder under "project-gantt.png"**](assets/project-gantt.png)

Back-up Plan
============

*This is part C) in the project plan [sv: projektredogörelse]*

References
==========

<!-- Do not edit by hand, references will be inserted and formatted automatically once you knit this document -->