### 2025-12-10 manuscript.qmd
Writing abstract + intro for manuscript

---
title: "Opportunities for improvement in the care of adult patients with traumatic brain injury"
format: html
author: "Vivek Kanwar"
bibliography: bibliography.bib
csl: vancouver.csl
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<!-- This is an R Markdown document. Markdown is a simple formatting
syntax for authoring HTML, PDF, and MS Word documents. For more
details on using R Markdown see http://rmarkdown.rstudio.com -->

<!-- This is a comment and will not be present in the compiled
document. Try "knitting" this document by clicking the knit button up
to the left, or by running `rmarkdown::render("manuscript.Rmd")` in
the console and you should only see headings -->

<!-- Structure and most comments are from the Strengthening the
reporting of observational studies in epidemiology (STROBE) statement,
see https://www.strobe-statement.org/. For more explanations see the
paper Strengthening the Reporting of Observational Studies in
Epidemiology (STROBE):explanation and elaboration.  -->

<!-- Indicate the study’s design with a commonly used term in the
title, for example: "Associations between resuscitation interventions
and opportunities for improvement in adult trauma patients: A cohort
study" -->

# Abstract
<!-- Provide in the abstract an informative and balanced summary of
what was done and what was found. Not more than 300 words. -->
## Background

Traumatic brain injury (TBI) is a major cause of death and disability after trauma. Trauma quality programmes increasingly use “opportunities for improvement” (OFI) to identify deviations from optimal care, but patterns of OFI in adult TBI are not well described. As such, this study aims to determine if and how often OFIs occur in the care of adult patients with TBI, describe their main characteristics, and examine whether patient- and process-related factors are associated with OFI

## Methods

A registry-based retrospective cohort study at Karolinska University Hospital in Solna, Sweden, centre using the local trauma registry (2012–2023) linked to the hospital’s trauma care quality database was conducted. Adult TBI patients were identified using the Abbreviated Injury Scale (AIS). The outcome was the presence of at least one OFI, classified as yes/no through a standardised multidisciplinary review. Patient- and process-related variables were analysed in complete-case logistic regression models with OFI status as the dependent variable.

## Results

OFIs were identified in 7.7% of TBI patients. In adjusted analyses, higher ISS (adjusted OR 1.05, 95% CI 1.03–1.07), higher Revised Trauma Score (adjusted OR 1.32, 95% CI 1.03–1.72; p = 0.034), admission to a high-dependency unit compared to the emergency ward (adjusted OR 3.50, 95% CI 1.30–11.13) and longer hospital stay (adjusted OR 1.02 per day, 95% CI 1.00–1.04) were independently associated with OFI.

## Conclusion

In this adult TBI cohort, OFIs were relatively uncommon but found in patients with high anatomical injury burden, seemingly stable physiology, intermediate care levels and longer hospital stays. These patterns suggest that trauma quality improvement efforts may be particularly relevant in TBI patients with complex care trajectories and those who appear physiologically stable despite substantial injury burden.

# Introduction

<!-- Explain the scientific background and rationale for the
investigation being reported. End by stating the aim or specific
objectives, including any prespecified hypotheses. The introduction
should not be longer than 1-1.5 pages, typically 4-5 paragraphs, each
6-8 lines long. Support every statement with a citation -->
Traumatic injury is a major global health problem and a leading cause of death and disability, particularly among people younger than 45 years [@Maerz2009]. Within this burden, traumatic brain injury (TBI) accounts for a large proportion of trauma-related mortality and long-term neurological impairment and has been highlighted as a priority area for trauma systems (2,3). TBI arises when external forces causes acute neurological impairment and initiate both primary structural damage and secondary pathophysiological cascades, including oedema, ischemia and metabolic disturbances which develop over time and critically influence long-term outcomes (4,5).

## Management of TBI

After initial stabilisation, management of TBI focuses on preventing secondary brain injury by securing the airway, ensuring adequate ventilation and rapidly correcting hypotension, as lapses in these areas are associated with increased mortality (6-9). In terms of imaging, rapid non-contrast head CT is essential to detect surgically liable lesions, while in patients with multi-system trauma, whole-body CT may be considered (10,11). International guidelines and reviews, including those from the Brain Trauma Foundation (BTF), provide recommendations on monitoring, control of intracranial- and cerebral perfusion pressure, and avoidance of potentially harmful interventions such as corticosteroids (5,7,13-15). Despite these frameworks, the heterogeneous and dynamic nature of TBI makes care particularly susceptible to delays, omissions and deviations from optimal practice (5,7,13).

## Trauma Quality Improvement

To address such vulnerabilities in trauma cohorts, trauma quality improvement programmes (TQIPs) systematically evaluate care processes and outcomes. These programmes use audit filters, structured case reviews and multidisciplinary morbidity and mortality conferences to find “opportunities for improvement” (OFI), defined as deviations from optimal care that may have contributed to adverse outcomes or  preventable harm (17-19). The World Health Organization and the American College of Surgeons provide guidance on how these programmes should be organised, emphasising continuous, multidisciplinary review of trauma cases and feedback to clinicians (18,19).

Previous studies in general trauma populations, and in mature trauma settings even, have shown that deviations and potentially preventable outcomes often arise from shortcomings in areas regarding, airway management, haemorrhage control, resuscitation and delays in imaging or missed injuries (3,20-23). Within this framework, patterns focused specifically on TBI is more limited. Studies of preventable or potentially preventable TBI deaths report recurrent issues such as delayed neurosurgical intervention, suboptimal early resuscitation and inadequate monitoring, suggesting that TBI may carry distinct vulnerabilities along the trauma pathway (24,25).

## Rationale for this study

TBI remains a challenge in modern system because even small deviations from optimal practice can markedly worsen outcomes. Although TQIPs aim to identify and correct such lapses, most research has focused on mixed trauma cohorts, leaving TBI-specific vulnerabilities under-characterised where also recent work suggests that patients with TBI remain at risk of OFIs (26). As such, this study aims to determine whether OFIs occur in the care of adult trauma patients with TBI, and, if so, to quantify their frequency, describe their characteristics, and examine whether patient- and process-related factors are associated with their occurrence.

# Methods

## Study design

<!-- Present key elements of study design -->

## Setting

<!-- Describe the setting, locations, and relevant dates, including
periods of recruitment, exposure, follow-up, and data collection -->

## Participants

<!-- Cohort study: Give the eligibility criteria, and the sources and
methods of selection of participants. Describe methods of
follow-up. For matched studies, give matching criteria and number of
exposed and unexposed -->

<!-- Case-control study: Give the eligibility criteria, and the
sources and + methods of case ascertainment and control
selection. Give the rationale for the choice of cases and
controls. For matched studies, give matching criteria and the number
of controls per case -->

<!-- Cross-sectional study: Give the eligibility criteria, and the
sources and methods of selection of participants -->

## Variables and data sources/measurements

<!-- Clearly define all outcomes, exposures, predictors, potential
confounders, and effect modifiers. Give diagnostic criteria, if
applicable. For each variable of interest, give sources of data and
details of methods of assessment (measurement). Describe comparability
of assessment methods if there is more than one group -->

## Bias

<!-- Describe any efforts to address potential sources of bias -->

## Study size

<!-- Explain how the study size was arrived at -->

## Quantitative variables

<!-- Explain how quantitative variables were handled in the
analyses. If applicable, describe which groupings were chosen and why
-->

## Statistical methods

<!--

(a) Describe all statistical methods, including those used to control
for confounding

(b) Describe any methods used to examine subgroups and interactions

(c) Explain how missing data were addressed

(d)

Cohort study: If applicable, explain how loss to follow-up was addressed

Case-control study:If applicable, explain how matching of cases and
controls was addressed

Cross-sectional study: If applicable, describe analytical methods
taking account of sampling strategy

(e) Describe any sensitivity analyses
-->

# Results

## Participants

<!-- (a) Report numbers of individuals at each stage of study—eg numbers
potentially eligible, examined for eligibility, confirmed eligible,
included in the study, completing follow-up, and analysed

(b) Give reasons for non-participation at each stage

(c) Consider use of a flow diagram -->

```{r flowchart, echo=FALSE}

library(glue)
library(DiagrammeRsvg)
source("main.R") ## This "imports" the main script file of your project and run any code in it

# Flowchart figure -----------------------------------------------------------

# ---- Counts for intermediate steps -----------------------------------------
n_excl_step1 <- n_total_merged_data - n_TBI_only_data
n_excl_step2 <- n_under18 + n_no_ofi

# ---- Labels ----------------------------------------------------------------
label_A <- glue(
  "Trauma registry 2012–2023\n",
  "n = {n_total_merged_data}"
)

label_B <- glue(
  "Identified TBI using AIS 1xxxxx.x\n",
  "n = {n_TBI_only_data}"
)

label_C <- glue(
  "Eligible adult TBI patients\n",
  "with OFI outcome\n",
  "n = {n_eligible}"
)

label_D <- glue(
  "Included (final analytic cohort)\n",
  "n = {n_final} adult TBI patients\n",
  "- OFI: {n_ofi} ({round(100 * n_ofi / n_final, 1)}%)\n",
  "- No OFI: {n_final - n_ofi}"
)

label_excl1 <- glue( 
  "Excluded (n = {n_excl_step1}):\n",
  "- No TBI according to AIS code"
  )

label_excl2 <- glue(
  "Excluded (n = {n_excl_step2}):\n",
  "- Age < 18 years: {n_under18}\n",
  "- No OFI outcome: {n_no_ofi}"
)

label_excl3 <- glue(
  "Excluded: missing data\n",
  "in covariates (n = {n_missing_cc})"
)

# ---- Producing the figure with exclusion boxes ------------------------------

g <- grViz(glue("
digraph TBI_flow {{

  graph [
    rankdir = TB,
    splines = ortho
  ]

  node [
    shape = box,
    style = rounded,
    fontname = 'Times New Roman',
    fontsize = 12,
    width = 3
  ]

  edge [
    arrowhead = normal
  ]

  // Main vertical line
  A [label = '{label_A}']
  B [label = '{label_B}']
  C [label = '{label_C}']
  D [label = '{label_D}']

  // Side exclusion boxes
  E1 [label = '{label_excl1}']
  E2 [label = '{label_excl2}']
  E3 [label = '{label_excl3}']

  // Invisible junction nodes on the vertical line
  J1 [shape=point, width=0.001, label=\"\", style=invis]
  J2 [shape=point, width=0.001, label=\"\", style=invis]
  J3 [shape=point, width=0.001, label=\"\", style=invis]

  // Main vertical line
  A -> J1 [arrowhead=none] 
  J1 -> B       
  B -> J2 [arrowhead=none] 
  J2 -> C 
  C -> J3 [arrowhead=none]
  J3 -> D

  // Branching arrows from the vertical line to exclusions
  J1 -> E1
  J2 -> E2
  J3 -> E3

  // Keep exclusions horizontally aligned with the junctions
  {{rank=same; J1; E1}}
  {{rank=same; J2; E2}}
  {{rank=same; J3; E3}}

}}
"))

# ---- Export (SVG + PNG) ----
svg_txt <- DiagrammeRsvg::export_svg(g)

out_svg <- file.path("assets", "flowchart.svg")
out_png <- file.path("assets", "flowchart.png")

cat(svg_txt, file = out_svg)
rsvg_png(out_svg, file = out_png, width = 1600, height = 2600)

```

![**Figure 1. Flowchart of patient selection for the analytical cohort**](assets/flowchart.png)

## Descriptive data

<!--  
(a) Give characteristics of study participants (eg demographic,
clinical, social) and information on exposures and potential
confounders

(b) Indicate number of participants with missing data for each
variable of interest

(c) Cohort study — Summarise follow-up time (eg, average and total
amount)
-->
```{r}
Descriptive.table1
```
## Outcome data

<!--  
Cohort study — Report numbers of outcome events or summary measures
over time

Case-control study — Report numbers in each exposure category, or
summary measures of exposure

Cross-sectional study — Report numbers of outcome events or summary
measures
-->

##  Main results

<!--  
(a) Give unadjusted estimates and, if applicable, confounder-adjusted
estimates and their precision (eg, 95% confidence interval). Make
clear which confounders were adjusted for and why they were included

(b) Report category boundaries when continuous variables were
categorized

(c) If relevant, consider translating estimates of relative risk into
absolute risk for a meaningful time period
-->

```{r}
Merged.Table
```

## Other analyses

<!--  
Report other analyses done—eg analyses of subgroups and interactions,
and sensitivity analyses 
-->

You can include code in this document like this:

```{r main, echo=FALSE}
source("main.R") ## This "imports" the main script file of your project and run any code in it
```

You can also embed plots:

```{r plot, echo=FALSE}
plot(pressure)
```

You can also mix text and code, so called inline code, like this: `r 2+5`.

# Discussion

## Key results

<!--
Summarise key results with reference to study objectives
-->

## Limitations

<!--
Discuss limitations of the study, taking into account sources of
potential bias or imprecision.  Discuss both direction and magnitude
of any potential bias
-->

## Interpretation

<!--
Give a cautious overall interpretation of results considering
objectives, limitations, multiplicity of analyses, results from
similar studies, and other relevant evidence
-->

## Generalisability

<!-- Discuss the generalisability (external validity) of the study results -->

# Conclusion

<!-- Give a short conclusion, corresponding to the aim -->

# References

<!-- Do not edit by hand, references will be inserted and formatted automatically once you knit this document -->